Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-01-15 TEBenchmarkPaper/article.tex
Include Alison's remarks in discussion

We do not claim that the use of such corpora exactly captures all interesting
conjectures of their definitions, or that those definitions exactly represent
all theories we may wish to explore. Rather, we consider our approach to offer a
pareto-optimal balance between theoretical rigour and experimental practicality,
at least in the short term.

If ATE matures to the point of providing \emph{truly novel} insights, even in
well-studied domains, our assumption breaks down because such results cannot (by
definition) appear in any existing corpus.

A fundamental limitation of our benchmarking approach is that it only applies to
systems which act in ``batch mode'', i.e. those which choose to halt after
emitting some output. Whilst all of the systems we have encountered are of this
form, an analogous benchmark may be desirable for systems which instead run
continuously or interactively (perhaps akin to McCarthy's
``advice taker''~\cite{McCarthy_Programs59}).

>>>>

Fundamentally, any mathematical reasoning system must decide on, and formalise,
what counts as ``the good'' in mathematics.  Obvious metrics such as ``true'' or
``provable'' include trivial tautologies, while at the same time failing to
capture the ``almost true'', which can be a valuable trigger for theory change,
as demonstrated by Lakatos in his case studies of mathematical development
\cite{lakatos}. ``Beautiful'' is another -- albeit vague -- commonly proposed
metric. Methods for finding out what that means have come from neuro-scientists,
such as the study by Zeki et al in which they test whether mathematicians'
experiences of abstract beauty correlates with the same brain activity as
experiences of sensory beauty \cite{zeki}; computer scientists, such as Colton's
metrics above which are proposed based largely on a computer scientist's
``intuition'' and argumentation about why a metric would be important (and
previous use of such metrics, and - in one isolated case (surprisingingness) -
on a single quote from a mathematician\footnote{Similarly, classification of a
MATHsAiD result as either interesting or uninteresting \cite{roy} was made by
the main system developer: in both cases the main developer had a background in
mathematics.}; and mathematicians themselves, such as Gowers' suggestion that we
can identify features which are commonly associated with good proofs
\cite{gowers}. All of these approaches rest upon the assumption that it makes
sense to speak of ``the good'' in mathematics. This assumption is supported by
concepts such as Erdos's ``The Book'' -- a theoretical construct of a book which
contains the most elegant proof of every theorem (brought to life in
\cite{Martin Aigner and Günter M. Ziegler Proofs from THE BOOK}). However,
empirical studies by psychologists such as Inglis call into question such
assumptions: work by Inglis and colleagues has shown that there is not a single
standard of validity among contemporary mathematicians %FIXME https://dspace.lboro.ac.uk/dspace-jspui/handle/2134/12150
nor beauty. %FIXME needs citation

We do not claim that our use of corpora as a ground truth exactly captures all
interesting conjectures of their definitions, or that those definitions exactly
represent all theories we may wish to explore. Rather, we consider our approach
to offer a pareto-optimal balance between theoretical rigour and experimental
practicality, at least in the short term. Futhermore, since research is already
on-going in these areas, we hope to at least improve on existing evaluation
practices and offer a common ground for future research.

Notably, if ATE matures to the point of providing \emph{truly novel} insights,
even in well-studied domains, then the assumptions we rely on break down because
such results cannot (by definition) appear in any existing corpus.

Another practical limitation limitation of our benchmarking approach is that it
only applies to systems which act in ``batch mode'', i.e. those which choose to
halt after emitting some output. Whilst all of the systems we have encountered
are of this form, an analogous benchmark may be desirable for systems which
instead run continuously or interactively (perhaps akin to McCarthy's
``advice taker''~\cite{McCarthy_Programs59}).
